---
title: "Hotel Data Week 4 Data Dive"
author: "Reshu Gupta"
date: "`r Sys.Date()`"
output:
  html_document: default
---
1. Initial setup and Configure the data set.</br>
2. Load the data set file in variable hotel_data files. <br>
3. Data set - Hotels : This data comes from an open hotel booking demand dataset from Antonio, Almeida and Nunes.<br>

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
# Load the 'dplyr' library
library(dplyr)
# Load the data into hotel_data for further use
hotel_data <- read.csv(file.choose())

```
In this section,<br>
1. Data set 'hotel_data' is summarized.<br>
2. then find the length of dataset - hotel_data by using nrow() and assign to variable - hotel_data_length.<br>
3. Calculate and print the size of subsample  (50% of hotel_data_lenght). <br>
```{r hotel_data}
#summery of hotel_data
summary(hotel_data)
#Calculate the length of dataset hotel_data
hotel_data_length<-nrow(hotel_data)
hotel_data_length
# Calculate the subsample size with 50% (according to question)
subsample_size <- round(hotel_data_length * 0.5)
# Print the samplesize
subsample_size
```
In this section, 5 sub samples have been created of size - subsample_size (subsample_size - 50% of hotel_data_length). <br>

```{r create_sub_sample_hotel_data}
# Create 5 random subsample of hotel data. We have used sample_n of dplyr for the same.
#hotel_data_subsample_df_1
#hotel_data_subsample_df_2
#hotel_data_subsample_df_3
#hotel_data_subsample_df_4
#hotel_data_subsample_df_5
hotel_data_subsample_df_1 <- hotel_data %>% sample_n(size = subsample_size, replace = TRUE)
hotel_data_subsample_df_2 <- hotel_data %>% sample_n(size = subsample_size, replace = TRUE)
hotel_data_subsample_df_3 <- hotel_data %>% sample_n(size = subsample_size, replace = TRUE)
hotel_data_subsample_df_4 <- hotel_data %>% sample_n(size = subsample_size, replace = TRUE)
hotel_data_subsample_df_5 <- hotel_data %>% sample_n(size = subsample_size, replace = TRUE)


```
Scrutinize #1

In this section, Ist Sub sample - hotel_data_subsample_df_1 has been scrutinize.<br>
1. Length of Sub sample - hotel_data_subsample_df_1. <br>
2. Print Ist five rows (for few column) of sub sample.<br>
3. Print Last five rows (for few column)  of sub sample. <br>
Note : By seeing 2 & 3, data consistency can be verified.<br>
4. Print the Internal structure of Sub sample.<br>
5. then summarized - sub sample.<br>

```{r scrutinize_sub_sample_hotel_data_1}
#Scrutinize these subsample hotel_data_subsample_df_1
# Total No of row in hotel_data_subsample_df_1
nrow(hotel_data_subsample_df_1)
# Top 5 Rows of hotel_data_subsample_df_1
#head(hotel_data_subsample_df_1)
head(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)

# Last 5 Rows of hotel_data_subsample_df_1
tail(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)

#  Internal structure of hotel_data_subsample_df_1
str(hotel_data_subsample_df_1)
#Summary of hotel_data_subsample_df_1
summary(hotel_data_subsample_df_1)

```
Scrutinize #2 <br>
In this section, IInd Sub sample - sub_sample_hotel_data_2 has been scrutinize.<br>
1. Length of Sub sample - hotel_data_subsample_df_1.<br>
2. Print Ist five rows (for few column)  of sub sample.<br>
3. Print Last five rows (for few column)  of sub sample. <br>
Note : By seeing 2 & 3, data consistency can be verified.<br>
4. Print the Internal structure of Sub sample.<br>
5. then summarized - sub sample  <br>


```{r scrutinize_sub_sample_hotel_data_2}
#Scrutinize these subsample hotel_data_subsample_df_2
# Total No of row in hotel_data_subsample_df_2
nrow(hotel_data_subsample_df_2)
# Top 5 Rows of hotel_data_subsample_df_2
head(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)
# Last 5 Rows of hotel_data_subsample_df_2
tail(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)

#  Internal structure of hotel_data_subsample_df_2
str(hotel_data_subsample_df_2)
#Summary of hotel_data_subsample_df_2
summary(hotel_data_subsample_df_2)

```
Scrutinize #3<br>
In this section, IIIrd Sub sample - sub_sample_hotel_data_3 has been scrutinize.<br>
1. Length of Sub sample - hotel_data_subsample_df_3.<br>
2. Print Ist five rows (for few column)  of sub sample.<br>
3. Print Last five rows (for few column)  of sub sample. <br>
Note : By seeing 2 & 3, data consistency can be verified.<br>
4. Print the Internal structure of Sub sample.<br>
5. then summarized - sub sample  <br>

```{r scrutinize_sub_sample_hotel_data_3}

#Scrutinize these subsample hotel_data_subsample_df_3
# Total No of row in hotel_data_subsample_df_3
nrow(hotel_data_subsample_df_3)
# Top 5 Rows of hotel_data_subsample_df_3
head(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)

# Last 5 Rows of hotel_data_subsample_df_3
tail(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)

#  Internal structure of hotel_data_subsample_df_3
str(hotel_data_subsample_df_3)
#Summary of hotel_data_subsample_df_3
summary(hotel_data_subsample_df_3)


```
Scrutinize #4<br>
In this section, IVth Sub sample - sub_sample_hotel_data_4 has been scrutinize.<br>
1. Length of Sub sample - hotel_data_subsample_df_4.<br>
2. Print Ist five rows (for few column)  of sub sample.<br>
3. Print Last five rows (for few column)  of sub sample. <br>
Note : By seeing 2 & 3, data consistency can be verified.<br>
4. Print the Internal structure of Sub sample.<br>
5. then summarized - sub sample  <br>

```{r scrutinize_sub_sample_hotel_data_4}

#Scrutinize these subsample hotel_data_subsample_df_4
# Total No of row in hotel_data_subsample_df_4
nrow(hotel_data_subsample_df_4)
# Top 5 Rows of hotel_data_subsample_df_4
head(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)
# Last 5 Rows of hotel_data_subsample_df_4
tail(hotel_data_subsample_df_1[,c("hotel","lead_time","meal","market_segment","distribution_channel","country")],5)

#  Internal structure of hotel_data_subsample_df_4
str(hotel_data_subsample_df_4)
#Summary of hotel_data_subsample_df_4
summary(hotel_data_subsample_df_4)

```
Scrutinize #5<br>
In this section, Vth Sub sample - sub_sample_hotel_data_5 has been scrutinize.<br>
1. Length of Sub sample - hotel_data_subsample_df_5.<br>
2. Print Ist five rows(for few column)  of sub sample.<br>
3. Print Last five rows(for few column)  of sub sample. <br>
Note : By seeing 2 & 3, data consistency can be verified.<br>
4. Print the Internal structure of Sub sample.<br>
5. then summarized - sub sample <br>


```{r  scrutinize_sub_sample_hotel_data_5}

#Scrutinize these subsample hotel_data_subsample_df_5
# Total No of row in hotel_data_subsample_df_5
nrow(hotel_data_subsample_df_5)
# Top 5 Rows of hotel_data_subsample_df_5
head(hotel_data_subsample_df_1[,c("hotel","meal","market_segment","distribution_channel","country")],5)
# Last 5 Rows of hotel_data_subsample_df_5
tail(hotel_data_subsample_df_1[,c("hotel","meal","market_segment","distribution_channel","country")],5)

#  Internal structure of hotel_data_subsample_df_5
str(hotel_data_subsample_df_5)
#Summary of hotel_data_subsample_df_5
summary(hotel_data_subsample_df_5)


```
## Box plot for a numeric variable -'lead_time'  in  subsample.



```{r pressure, echo=TRUE}
boxplot(hotel_data_subsample_df_1$lead_time, hotel_data_subsample_df_2$lead_time, hotel_data_subsample_df_3$lead_time, hotel_data_subsample_df_4$lead_time, hotel_data_subsample_df_5$lead_time, names = c("1", "2", "3", "4", "5"), main = "Lead Time Distribution by Sub Sample 1", ylab = "Lead Time")

```
<br>
#Anomaly Detection for sub sample data frame#1 : - hotel_data_subsample_df_1

```{r  anomaly_hotel_data_subsample_df_1}
hotel_data_subsample_df_1 %>%
  group_by(hotel) %>%
  summarise(count = n())

```
<br>
#Anomaly Detection for data frame#2 : - hotel_data_subsample_df_2
```{r  anomaly_hotel_data_subsample_df_2}
hotel_data_subsample_df_2 %>%
  group_by(hotel) %>%
  summarise(count = n())
```
#Anomaly Detection for data frame#3 : - hotel_data_subsample_df_3

```{r anomaly_hotel_data_subsample_df_3}
hotel_data_subsample_df_3 %>%
  group_by(hotel) %>%
  summarise(count = n())

```

#Anomaly Detection for data frame#4 : - hotel_data_subsample_df_4

```{r anomaly_hotel_data_subsample_df_4}
hotel_data_subsample_df_4 %>%
  group_by(hotel) %>%
  summarise(count = n())

```
#Anomaly Detection for data frame#5 : - hotel_data_subsample_df_5

```{r anomaly_hotel_data_subsample_df_5}
hotel_data_subsample_df_5 %>%
  group_by(hotel) %>%
  summarise(count = n())

```

# Consistency Analysis on data frame #1

```{r hotel_data_subsample_df_1_consistency_analysis}
table(hotel_data_subsample_df_1$country)

```
# Consistency Analysis on data frame #2

```{r hotel_data_subsample_df_2_consistency_analysis}
table(hotel_data_subsample_df_2$country)

```

# Consistency Analysis on data frame #3

```{r hotel_data_subsample_df_3_consistency_analysis}
table(hotel_data_subsample_df_3$country)

```

# Consistency Analysis on data frame #4

```{r hotel_data_subsample_df_4_consistency_analysis}
table(hotel_data_subsample_df_4$country)

```



# Consistency Analysis on data frame #5

```{r hotel_data_subsample_df_5_consistency_analysis}
table(hotel_data_subsample_df_5$country)

```


#Monte Carlo Simulations of data set hotel_data_subsample_df_1

```{r hotel_data_df_monte_carlo_sim}
monte_carlo_sim <- rnorm(2000, mean = mean(hotel_data_subsample_df_1$lead_time), sd = sd(hotel_data$lead_time))

# Plot (histograms) the distribution of the Monte Carlo simulation
hist(monte_carlo_sim, main = "Monte Carlo Simulation of Lead Time", xlab = "Lead Time")



```

##<b>Monte Carlo Simulations</b>
##<p style='font-size:normal'>Monte Carlo Simulations histograms diagram is graphical representationof frequency distribution of data set which were based on randam numbers genetated using Monte Carlo simulations.</p>
##<p> In above histograms diagram, The Title of Graph is - "Monte Carlo Simulation of Lead Time" which is indicating that the plot represt the distribution of Lead Times(on X Axis) derived from Monte Carlo simulation. The X-axis label - Lead Time specifies the variable being measured. </p>
##<p>The hight of each bar represnts the frequency of lead times falling within a specific range. Higher Bar means a higher occurence of lead time with in the range.</p>
##<p>The histogram can be compared to expected or historical lead time data. Discrepancies may highlight areas where the simulation differs from real-world observations, prompting further investigation or refinement of the simulation model.</p>
##<p>The insights gained from the histogram can be used for decision support, especially in scenarios where lead time variability is a critical factor. Understanding the distribution helps in making informed decisions and developing strategies to manage lead time uncertainties.</p>

<p>Thank You.!!!</P>

